\chapter{Preliminaries.}
We start this chapter reminding the basic definitions and theorems in topology and measure theory, since they are needed to have a suitable framework to discuss the optimal transport problem and its applications.

 
\section{Definitions and important theorems to remember.}

\subsection{Topology.}
We start with the definition of topology that is needed to introduce a notion of continuity. We refer to \cite{munkres2000topology} for more details in topology.

\begin{definition}[Topology]
	A topology on a set $X$ is a collection $\mathcal T$ of subsets of $X$ having the following properties
	\begin{itemize}
		\item  The space $X$ itself and $\emptyset$ are in $\mathcal T$.
		\item  The union of the elements of any sub-collection of $\mathcal T$ is in $\mathcal T$.
		\item The intersection of the elements of any finite sub-collection of $\mathcal T$ is in $\mathcal T$.
	\end{itemize}
\end{definition}

A pair $\parentheses{X, \mathcal{T}}$ is called a topological space. The elements of $\mathcal T$ are called open sets. The complements of the sets of $\mathcal T$ are called closed sets. The interior of a set $A$, is defined as the biggest open set contained in $A$. Similarly, the closure of a set $A$, is defined as the smallest closed set containing $A$. We use indistinctly the notation $\inter(A)$ and $A^{\circ}$ for the interior of a set $A$. In the same way, for the closure we use the notation $\closu(A)$ or $\bar{A}$. An equivalent way to define the same ideas is give by the following,

\begin{align*}
	\inter(A)=\bigcup_{B \text{ is open.}} B \\
	\closu(A)=\bigcap_{B \text{ is closed.}} B
\end{align*}

We remark that a set is open if and only if $A=A^{\circ}$, and a set is closed if and only if $A=\bar{A}$.  


We call a neighborhood of $x$ an element of $\mathcal{T}$ containing $x$.
\begin{definition}[Topological Basis.]
	Give a set $X$ endowed with a topology $\mathcal T$. We call a basis for $\mathcal{T}$ is a collection $\mathcal{B}$ of subsets of $X$ (called basis elements), such that,
	\begin{enumerate}
		\item  For each $x\in X$, there is at least one basis element $B$ containing $x$. 
		\item  If $x$ belongs to the intersection of two basis elements $B_1$ and $B_2$, then there is a basis element $B_3$ containing $x$ such that $B_3 \in B_1 \cap B_2$. 
	\end{enumerate}
\end{definition}



\begin{definition}[Dense set]
	A subset $D$ of a topological space $X$ is dense in $X$ if for any point $x$ in $X$, any neighborhood of $x$ contains at least one point from $D$.
\end{definition}

\begin{definition}[Separable space]
	A topological space is called separable if it contains a countable, dense subset.
\end{definition}

Topologies in which one element is not a closed set, or in which a sequence can converge to more than one point, are not really interesting for practical problems. If such things are allowed the theorems that one can prove are limited. A mathematician Felix Hausdorff suggested to add the following condition:
  
\begin{definition}[Hausdorff space]
A topological space $X$ is called a Hausdorff space if for each pair $x_1$, $x_2$ of distinct points of $X$, there exist neighborhoods $U_1$, and $U_2$ of $x_1$ and $x_2$, respectively, that are disjoint. 
\end{definition}

\begin{definition}[Distance]
\end{definition}
\begin{definition}[Metric Spaces]
	
\end{definition}

\begin{definition}[Completeness]
	A metric space $X$ is called complete if every Cauchy-Sequence of points in $X$ has a limit that is also in $X$. 
\end{definition}

\begin{definition}[Completely metrizable space]
	
\end{definition}
There is a subtle difference between complete metric space and completely metrizable space.  And the difference lies on the words ``\textit{there exists at least a metric...}" in the completely metrizable definition, and ``\textit{given a metric}''. Complete metrizable is a topological property while completeness is a property of the chosen metric.
\begin{definition}[Polish space]
	We call Polish space to any topological space that is separable and completely metrizable.
\end{definition}

\begin{definition}[Sequentially compact]
	A subset $K$ of a metric space $X$ is said to be compact if from any sequence $x_n$, we can extract a converging subsequence $x_{n_k} \rightarrow x \in K$.
\end{definition}

\begin{definition}[Compactness]
		A subset $K$ of a metric space $X$ is compact if every open cover of $K$ has a finite subcover. 
\end{definition}

\begin{theorem}
	A subset of a metric space is compact if and only if it is sequentially compact.
\end{theorem}


\begin{definition}[Liminf and Limsup]
	Let $X$ be a Hausdorff space. Let $\mathcal{V}(x_0)$ be a topological basis of $X$, such that all $V\in \mathcal{V}$ contains $x_0$. Let $f: X\rightarrow \Realex$ a functional valued in $\Realex$. We define, 
	\begin{equation*}
	\liminf_{x\rightarrow x_0} f(x)= \sup_{V\in \mathcal{V}(x_0)} \inf_{s\in V}\ f(s)
	\end{equation*}
	\begin{equation*}
	\limsup_{x\rightarrow x_0} f(x)= \inf_{V\in \mathcal{V}(x_0)} \sup_{s\in V}\ f(s)
	\end{equation*}
	The above definitions can be expressed in terms of sequences of real numbers. Let $(x_n)_{n \in \Naturals}$ be a sequence in $X$, the above formulation is equivalent to say.
	\begin{equation*}
	\liminf_{n\in \Naturals} x_n := \lim_{n\rightarrow \infty} \parentheses{\inf_{m\geq n} x_m}
	\end{equation*}
	Equivalently for $\limsup$,
	\begin{equation*}
	\liminf_{n\in \Naturals} x_n := \lim_{n\rightarrow \infty} \parentheses{\sup_{m\geq n} x_m}
	\end{equation*}
	Please note that the convergence to some point $x_0$, $(x_n)_{n\in \Naturals}\rightarrow x_0$ is not required in the last definitions.
\end{definition}


\subsection{Functional Analysis.}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%Functional Analysis %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%




\begin{definition}[Linear Space]
\end{definition}

\begin{definition}[Banach Space]
\end{definition}

\begin{definition}[Inner product]
\end{definition}

\begin{definition}[Hilbert Space]
\end{definition}

\begin{definition}[Continuity]
	
\end{definition}


For a given be a metric space $X$ . We denote the set of continuous, real-valued functions $f:X\rightarrow \Real$ by $C(X)$.

\begin{theorem}
	Let $K\subset X$ a compact subset of a metric space $X$. The space $C(K)$ is complete.
\end{theorem}
A natural norm on spaces of continuous functions is the uniform norm (also called infinity norm), which is defined by,
\begin{equation*}
\norm{f}_{\infty}=\sup_{x\in X} \abs{f(x)}
\end{equation*}
The norm $\norm{f}_{\infty}$ is finite if and only if $f$ is bounded. And we use $C_b(X)$ to refer the space of bounded functions on $X$.

\begin{definition}
	Let $f$ be a real valued function, $f:X\rightarrow \Real$ on a metric space. The \textbf{support of a function}, $\supp f$ is the closure of the set on which $f$ is nonzero.
	\begin{equation*}
	\supp f = \closu\parentheses{{\braces{x\in X: \quad f(x)\neq 0}}}
	\end{equation*}
	
\end{definition}
We say that $f$ has compact support if $\supp f$ is a compact subset of $X$, and denote the space of continuous functions on $X$ with compact support by $C_c(X)$.

The space $C_c(X)$ is a linear subspace of $C_b(X)$, but it does not need to be closed. 

\begin{definition}
	Suppose that X is a separable and locally compact metric space.	We say that a real valued function $f$ belongs to $C_0(X)$ if and only if $f\in C(X)$, and for every $\epsilon >0$, there exists a compact set $K\subset X$ such that $\abs{f}<\epsilon$ on $X\backslash K$. 
\end{definition}

\begin{definition}
	Let $\mathcal F$ be a family of functions from a metric space $\parentheses{X, d}$ to a metric space $\parentheses{Y, d}$. The family $\mathcal{F}$ is equicontinuous if for every $x\in X$ and $\epsilon > 0$ there is $\delta >0$ such that $d(x,y)<\delta$ implies $d(f(x), f(y))<\epsilon$ for all $f\in \mathcal F$.
\end{definition}

\begin{theorem}
	An equicontinuous family of functions from a compact metric space to a metric space is uniformly equicontinuous. 
\end{theorem}

\begin{theorem}[Ascoli-Arzel\`a]
	Let $K$ be a compact metric space. A subset $M$ of the set of continuous functions $M\subset C(K)$ is compact if and only if it is closed, bounded and equicontinuous. That is any sequence $(f_n)_{n\in\Naturals}$ in $M$ admits a subsequence converging to $f$ in $M$.
\end{theorem} 


\begin{definition}[Proper convex function]
	Let $f: X\rightarrow\Realex$, a function taking values in the extended real number line. We call it proper convex function if $\exists x \in X$ such that $f(x)<\infty$. And $\forall x \in X$, $f(x)>-\infty$.
\end{definition}

\begin{definition}[Projection of a Cartesian Product.]
	Let $\proj_x: X\times Y \rightarrow X$ be defined by the equation
	\begin{equation*}
		\proj_x(x,y)= x;
	\end{equation*}
	
	Equivalently, let $\proj_y: X\times Y \rightarrow Y$ be defined by,
	\begin{equation*}
		\proj_y(x,y)=y
	\end{equation*}
	These maps $\proj_x$ and $\proj_y$ are called the projections of $X\times Y$ onto $X$ and $Y$ respectively.	
\end{definition}

We can generalize a definition over a general Cartesian product. Given a set $X$, we define a \textit{$J$-tuple} of elements of $X$ to be a function $\mathbf{x}:J\rightarrow X$. If $\alpha$ is an element of $J$, we often denote the value of $\mathbf{x}$ at $\alpha$ by $x_\alpha$ rather than $\mathbf{x}(\alpha)$; we call it the $\alpha$-th coordinate of $\mathbf{x}$. And we often denote the function $\mathbf{x}$ by the symbol.

\begin{equation*}
	(x_\alpha)_{\alpha \in J}
\end{equation*} 

Let $\braces{A_\alpha}_{\alpha \in J}$ be a set of indexed family of sets; let $X=\bigcup_{\alpha \in J}A_{\alpha}$. The \textit{Cartesian product} of this indexed family, denoted by
\begin{equation*}
	\prod_{\alpha \in J}A_\alpha
\end{equation*}	
is defined to be the set of all $J$-tuples $\parentheses{x_\alpha}_{\alpha \in J}$ of elements of $X$ such that $x_\alpha \in A_\alpha$ for each $\alpha \in J$. That is, it is the set of all functions,
\begin{equation*}
\mathbf{x}: J\rightarrow \bigcup_{\alpha \in J} A_\alpha
\end{equation*}
such that $\mathbf{x}(\alpha)\in A_\alpha$ for each $\alpha \in J$.


\begin{definition}[Lower Semicontinuity]
	On a complete metric space $X$, a function $f: X\rightarrow\Realex$ is said to be lower semi-continuous (l.s.c.) if for every sequence $\parentheses{x_n}_{n\in\Naturals}$ converging to $x\in X$, we have 
	\begin{equation*}
	f(x)\leq \liminf_{n\in \Naturals} f(x_n)
	\end{equation*}
\end{definition}

We can see from the above definition that any continuous function is lower-semicontinuous. In other words, lower-semicontinuity is a milder requirement than continuity, although it preserves interesting properties that can be exploited in optimization.  

\begin{proposition}
	Let $f: X \rightarrow \Realex$ be a convex and lower-semicontinuous function.
	Assume that there exists $x_0 \in X$ such that $f(x_0) = -\infty$. Then $f$ is nowhere finite on $X$.
\end{proposition} 


\begin{theorem} If $ f_\alpha $ is an arbitrary family of lower semi-continuous functions on $X$, then $f=\sup_\alpha f_\alpha$ is also lower-semicontinuous.
\end{theorem}


\begin{definition}[Lipschitz condition]

\end{definition}

\begin{theorem}
	Let $f:X\rightarrow \Realex\backslash{-\infty}$ be a function bounded from below. Then $f$ is l.s.c. if and only if there exists a sequence $(f_n)_{n\in \Naturals}$ of $k$-Lipschitz functions such that for every $x\in X$, $f_n(x)$ converges increasingly to $f(x)$.
\end{theorem}

\begin{definition}
	Let $F:X\rightarrow \Realex$ be a given functional bounded from below on a metric space $X$.  Let $\mathcal G$ be the set of lower semicontinuous functions $G: X\rightarrow \Realex$, such that $G\leq F$. We call a relaxation the supremum of $\mathcal G$.  This functional does exist since the supremum of an arbitrary family of lower semicontinuous functions is also lower semicontinuous. It is possible to have a representation formula as follows:
	\begin{equation}
	\bar F(x) = \inf \braces{\liminf_{n\in \Naturals} F(x_n):\ x_n\rightarrow x}.
	\end{equation}
	As consequence of this definition we see that $F \geq \bar F$ implies $\inf F \geq \inf \bar F$. Let $l=\inf F$ then $F\geq l$. A constant function is lower semicontinuous. Therefore, $\bar F \geq l$ and $\inf \bar F \geq \inf F$. Implying that the infimum of both $F$ and its regularization $\bar F$ coincide, i.e. $\inf \bar F = \inf F$. 
\end{definition}


\begin{theorem}[Maxima and Minima]
	Let $X$ be a compact metric space and $f: X\rightarrow \Real$ is continuous, real-valued function. Then $f$ is bounded on $X$ and attains its maximum and minimum. That is, there are $x$, $y$ belonging to $X$ such that,
	\begin{equation*}
	f(x)=\inf_{z\in X}f(z) \qquad \mand \qquad f(y)=\sup_{z \in X} f(z)
	\end{equation*}
\end{theorem}

Continuity is a strong requirement. Luckily, we can assure the existence of a minimizer of lower-semicontinuous functionals (or maximizer for upper-semicontinuity). The usual procedure to prove existence of a minimizer is making use of Weierstrass' criterion. We take a minimizing sequence and then we prove that the space in which we are trying to find a minimizer element is compact.

\begin{theorem}[Weierstrass' criterion for existence of minimizers]
	If $f: X\rightarrow\Realex$ is lower semi-continuous and $X$ is compact, then there exists $\hat x \in X$ such $f(\hat x)=\min\braces{f(x):\ x\in X}$.
	\begin{proof}
		Define $l:=\inf\braces{f(x): \, x \in X} \in \Realex$, notice that $l=+\infty$ only if $f$ is identically $+\infty$, then this case is trivial since any point minimizes $f$. By compactness there exists a minimizing sequence $x_n$, that is $f(x_n)\rightarrow l$. By compactness we can extract a subsequence converging to some $\hat x$ such that $\hat x \in X$. By lower-semicontinuity of $f$, we have that $f(\hat x)\liminf_n f(x_n)=l$. Since $l$ is the infimum $l\leq f(\hat x)$. This proves that $l=f(\hat{x}) \in \Real$.
	\end{proof}
\end{theorem}

We can apply the above analysis using a notion of upper-semicontinuity and compactness to find the maximum.

\begin{definition}[Topological Dual]
	If $X$ is a normed space, the dual space $\asterisk{X}=\mathcal B(X, \Real)$. Consists of all linear and bounded functionals mapping from $X$ to $\Real$.
\end{definition}

\begin{definition}[Weak compactness in dual spaces]
	A sequence $x_n$ in a Banach space $X$ is said to be weakly converging to $x$, and we write $x_n \rightharpoonup x$, if for every $\xi \in \asterisk{X}$. We have $\anglesbox{\xi, x_n}\rightarrow \anglesbox{\xi, x}$. A sequence $\xi_n \in \asterisk{X}$ is said to be weakly-* converging to $\xi$, and we write $\xi_n \rightharpoonup \xi$, if for every $x\in X$ we have $\anglesbox{\xi_n, x} \rightarrow \anglesbox{\xi, n}$.
\end{definition}

\begin{theorem}[Banach-Alaouglu]
	If $X$ is separable and $\xi_n$ is a bounded sequence in $\asterisk{X}$, then there exists a subsequence $\xi_{n_k}$ weakly converging to some $\xi \in \asterisk{X}$.
\end{theorem}

The Banach-Alaouglu's theorem is a well known result in functional analysis, an equivalent formulation is saying the closed unit ball in $\asterisk{X}$ is weak-* compact.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% Measure Theory %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Measure Theory}

We refer \cite{bogachev2007Measure}
\begin{definition}[Sigma Algebra]
	
	An algebra of sets $\mathcal{A}$ is a class of subsets of some fixed set $X$ (called the space) such that
	
	\begin{itemize}
		\item $X$ and $\emptyset$ belong to $\mathcal{A}$.
		\item If $A, B \in \mathcal{A}$, then $A\cap B \in \mathcal{A}$, $A\cup B \in \mathcal A$, $A\backslash B \in \mathcal A$.
	\end{itemize}
	
	An algebra of sets is a $\sigma$-algebra if for any sequence of sets $A_n \in \mathcal A$ we have $\mathcal{A}\ni \bigcup_{n\in \Naturals}  A_n $.
\end{definition}

\begin{definition}[Measure Space] 
	A pair $(X, \mathcal A)$ consisting of a set $X$ and a $\sigma$-algebra $\mathcal A$ of its subsets is called a measurable space.
\end{definition}

\begin{definition}[Borel $\sigma$-algebra]
	The Borel $\sigma$-algebra $\mathcal{B}(\Real^n)$ of $\Real^n$ is the $\sigma$-algebra generated by all open sets. The sets in a Borel $\sigma$-algebra are called Borel sets. For any set $E\subset \Real^n$, let $\mathcal B(E)$ denote the class of all sets of the form $E\cap B$, where $B\in\mathcal B(\Real^n)$.
\end{definition}

Given a topological space, in this text consider the Borel $\sigma$-algebra unless stated otherwise. 

\begin{definition}[Measure]
	A real-valued set function $\mu:\mathcal{A}\rightarrow \Realex$ on a class of sets $\mathcal A$ is called countably additive if 
	\begin{equation*}
		\mu\parentheses{\bigcup_{n=1}^{\infty} A_n}=\sum_{n=1}^{\infty}\mu(A_n)
	\end{equation*}
	for all pairwise disjoint sets $A_n$ in $\mathcal A$ such that $\mathcal A \ni \bigcup_{n=1}^{\infty}A_i$. A countably additive set function defined on an algebra is called a measure.
\end{definition}

Given a measure space $X$, we denote $\MeasureSp(X)$ and $\PMeasureSp(X)$ to refer the set of finite measures and positive finite measures on $X$, respectively. 

\begin{definition}
	A countably additive measure $\mu$ on a $\sigma$-algebra of subsets of a space $X$ is called a probability measure if $\mu\geq 0$ and $\mu(X)=1$.
	A triple $(X, \mathcal A, \mu)$ is called a measure space if $\mu$ is a
	nonnegative measure on a $\sigma$-algebra A of subset of a set X. If $\mu$ is a probability measure, then $(X, \mathcal A, \mu)$ is called a probability space. 
\end{definition}


\begin{definition}[Lebesgue Measure]
\end{definition}

\begin{definition}[Hausdorff Measure]
\end{definition}

\begin{definition}[Probability]
	A countably additive measure $\mu$ on a $\sigma$-algebra of subsets of a space $X$ is called a probability measure if $\mu \geq 0$ and $\mu(X)= 1$.
	
	A triple $(X, \mathcal A, \mu )$ is called a measure space if $\mu$ is a
	nonnegative measure on a $\sigma$-algebra $\mathcal A$ of subset of a set $X$. If $\mu$ is a probability measure, then $(X, \mathcal A, \mu)$ is called a probability space.
\end{definition}

\begin{definition}[Product $\sigma$-algebra ]
Let $\left(X_1 , \mathcal{A}_1 , \mu_1 \right)$ and $\left(X_1 , \mathcal{A}_1 , \mu_1 \right)$ be two spaces with finite nonnegative measures. On the space $X_1\times X_2$ we consider sets of the form $A_1\times A_2$, where $A_i \in \mathcal{A}_i$, called measurable rectangles. Let $\mu_1\times \mu_2 \parentheses{A_1\times A_2}:= \mu_1(A_1)\mu_2(A_2)$. 

Let $A_1\otimes A_2$ denote the $\sigma$-algebra generated by all measurable rectangles; this $\sigma$-algebra is called the product of the $\sigma$-algebras $A_1$ and $A_2$.  

\end{definition}

\begin{theorem}
The set function $\mu_1\times \mu_2$ is countably additive on the algebra generated by all measurable rectangles and uniquely extends to a countably additive measure, denoted by $\mu_1\otimes \mu_2$.
\end{theorem}

\begin{definition}[Image Measure]
	Let $\parentheses{X, \mathcal{A}_X}$  and $\parentheses{Y, \mathcal{A}_Y}$ be two measurable spaces. Let $T:X\rightarrow Y$ be a measurable map from $X$ to $Y$. Let $\mu$ be a measure $\mu: \mathcal{A}_X\rightarrow \Realex_+$, then the image measure (or pushforward measure) $\Tmeasure{\mu}{T}:\mathcal{A}_2\rightarrow \Realex_+$ is given by,
	\begin{equation*}
		\Tmeasure{\mu}{T}(B) = \mu\parentheses{T^-1(B)}, \quad \forall B \in \mathcal{A}_Y.
	\end{equation*}
\end{definition}


\begin{theorem}
	Let $\parentheses{X, \mathcal{A}_X}$  and $\parentheses{Y, \mathcal{A}_Y}$ be two measurable spaces. Let $\mu$ be a nonnegative measure. A $\mathcal{A}_2$-measurable function $g$ on $Y$ is integrable with respect the measure $\mu\circ f^{-1}$ precisely when the function $g\circ f$ is integrable with respect to $\mu$. In addition we have, 
	\begin{equation*}
		\int_Y g(y)\mu\circ f^{-1}(\dy) = \int_X g(f(x))\mu(\dx)
	\end{equation*}
\end{theorem}

The space of Borel probability measures on $X$ is denoted by $\PlanSp(X)$. The weak topology on $\PlanSp(X)$ is induced by convergence against bounded continuous test functions on $X$, that is $C_b(X)$.

\begin{definition}[Atom and atomless measures]
	The set $A \in \mathcal A$ is called an atom of the measure $\mu$
	if $\mu(A) > 0$ and every set $B \subset A$ from $\mathcal A$ has measure either $0$ or $\mu(A)$. If there are no atoms, then the measure $\mu$ is called atomless. 
\end{definition}

A measure over a set $\Omega\subset \Real$ is atomless if $\forall x \in \Omega$, we have  $\mu(\{x\})=0$. The Dirac's measure is not atomless. 
\begin{definition}[Absolutely continuity and singularity]
	Let $\mu$ and $\nu$ be countably additive measures on a measurable space $\parentheses{X, \mathcal A}$. 
	\begin{itemize}
		\item The measure $\nu$ is called absolutely continuous with respect to $\mu$ if $\abs{\nu}(A)=0$ for every set $A$ with $\abs{\mu}=0$. We use the notation $\nu \ll \mu$.
		\item The measure $\nu$ is called singular with respect to $\mu$ if there exists a set $A\in \mathcal{A}$ such that
		\begin{equation*}
			\abs{\mu}(A)=0 \quad \text{ and }\quad  \abs{v}(X\backslash A)=0
		\end{equation*}
	\end{itemize}
	
	If $\nu \ll \mu $ and $\mu \ll \nu$, then the measures $\mu$ and $\nu$ are equivalent. We use the notation $\mu \sim \nu$ to refer this situation.
\end{definition}
The above definition allows us to introduce the Radon-Nikodym theorem that is one of the main results in measure theory.
\begin{theorem}[Radon–Nikodym theorem]
	Let $\mu$ and $\nu$ be two finite measures on a space $(X, \mathcal A)$.
	The measure $\nu$ is absolutely continuous with respect to the measure $\mu$ precisely	when there exists a $\mu$-integrable function $f$ such that $\nu$ is given by
	\begin{equation*}
		v(A)=\int_A f\dmu
	\end{equation*}
\end{theorem}

\begin{definition}[$L^p$ Spaces]
	
\end{definition}


\begin{definition}
	A sequence $\mu_n$ probability measures over $X$ is said to be tight if for every $\epsilon> 0$, there exists a compact subset $K\subset X$ such that $\mu_n\parentheses{X\backslash K}< \epsilon$ for every $n$.
\end{definition}




\begin{theorem}[Lebesgue dominated convergence theorem]
	Suppose that $\mu$-integrable functions $f_n$ converge almost everywhere to a function $f$. If there exists a $\mu$-integrable function $\Phi$ such that,
	\begin{equation*}
		\abs{f_n}(x)\leq \Phi (x), \quad \text{almost everywhere for every } n
	\end{equation*}
	then the function is integrable and
	\begin{equation*}
		\int_X f(x)\dmu(x) = \lim_{n\rightarrow\infty} \int_X f_n(x)\dmu(x)
	\end{equation*}
	In addition,
	\begin{equation*}
		\lim_{n\rightarrow \infty } \int \abs{f(x)-f_n(x)}\dmu(x) =0
	\end{equation*}
\end{theorem}

\begin{theorem}[Monotone Convergence]
	Let $(f_n)_{n\in \Naturals}$ be a sequence of $\mu$-integrable functions such that $f_n(x) \leq f_{n+1}$ almost everywhere for each $n \in \Naturals$. Suppose that
	\begin{equation}
	\sup_{n\in \Naturals}\int_X f_n(x)\dmu(x) < \infty
	\end{equation}
	
	Then the function $f(x)=\lim_{n\rightarrow\infty} f_n(x)$ is almost everywhere finite and integrable. In addition the following equality holds,
	\begin{equation*}
		\int_X f(x)\dmu(x) = \lim_{n\rightarrow\infty} \int_X f_n(x)\dmu(x)	
	\end{equation*}
\end{theorem}

\begin{theorem}[Fatou's Theorem]
	Let $(f_n)_{n\in \Naturals}$ be a sequence of nonnegative $\mu$-integrable functions convergent to a function $f$ almost everywhere and let
	\begin{equation*}
		\sup_{n\in \Naturals} \int_X f_n(x)\dmu(x)\leq K <\infty
	\end{equation*}
	
	Then, the function $f$ is $\mu$-integrable and 
	\begin{equation*}
		\int_X f(x)\dmu(x) \leq K
	\end{equation*}
	
	Moreover, 
	\begin{equation*}
		\int_X f(x)\dmu(x)\leq \liminf_{n\rightarrow\infty} \int_{X} f_n (x)\dmu(x)
	\end{equation*}
\end{theorem}

\begin{corollary}
	Let $\parentheses{f_n}_{n\in \Naturals}$ be a sequence of nonnegative $\mu$-integrable functions such that 
	\begin{equation*}
		\sup_{n\in \Naturals}\int_{X} f_n(x)\dmu(x) \leq K < \infty
	\end{equation*}
	Then the function $\liminf_{n\rightarrow \infty} f_n$ is $\mu$-integrable and one has 
	\begin{equation}
		\int_X \liminf_{n\rightarrow\infty}f_n(x)\dmu(x)\leq \liminf_{n\rightarrow \infty} \int_X f_n(x)\dmu(x) \leq K
	\end{equation}
\end{corollary}

\begin{corollary}
The dominated convergence theorem and Fatou’s theorem remain valid if in place of almost everywhere convergence in their hypotheses we require convergence of $(f_n)_{n\in \Naturals}$ to $f$ in measure $\mu$.
\end{corollary}



\begin{definition}[Tightness] Let $(X, \mathcal T)$ a topological space, and let $\mathcal A$ a $\sigma$-algebra on $X$ that contains the topology $\mathcal{T}$. Let $M$ be a collection of measures defined on $\mathcal{A}$. The collection $M$ is called tight  if for every $\epsilon>0$ there is a compact subset $K_\epsilon$ of $X$ such that, for all measures $\mu\in M$ we have,
	\begin{equation*}
	\abs{\mu}\parentheses{X \backslash K_\epsilon} < \epsilon
	\end{equation*} 
\end{definition}


\begin{theorem}[Prokhorov]
	Suppose that $\mu_n $ is a tight sequence of probability measures over a Polish space $X$. Then there exists $\mu\in\PlanSp(X)$ and a subsequence $\mu_{n_k}$ such that $\mu_{n_k}\rightharpoonup \mu $, in duality with $C_b(X)$. Conversely, every sequence $\mu_{n_k}\rightharpoonup \mu$ is tight.
\end{theorem}

\begin{definition}
	Let $\parentheses{X, \mathcal A, \mu}$ be a probability space. Then every Borel-measurable mapping $\X: X\rightarrow \Real$ with for all $B\in \mathcal B(\Real) $ is a random variable, denoted by $\X: \parentheses{X, \mathcal A}\rightarrow \parentheses{\Real, \mathcal B (\Real)} $
\end{definition}

\begin{definition}[Duality between $C_0$ and $\MeasureSp$ ]
		
\end{definition}


\textbf{Explanation about notions of convergence with bounded functionals and vanishing in infinity functions}. If $X$ is compact we have $C_0(X)=C_b(X)=C(X)$ if $X$ and both notions of convergence coincide.



\begin{theorem}[Rademacher]
	Let $f:\Real^d\rightarrow\Real$ be a Lipschitz continuous function. Then the set of points where $f$ is not differentiable is negligible for the Lebesgue measure.
\end{theorem}


\begin{lemma}
If $\mu$, $\nu$ are two probability measures on the real line $\Real$ and $\mu$ is atomless, then there exists at least a transport map $T$ such that $\Tmeasure{\mu}{T}=\nu$.
\end{lemma}

\begin{lemma}
There exists a Borel map $\sigma_d: \Real^d\rightarrow \Real$ which is injective, its image is a Borel subset of $\Real$, and its inverse map is Borel measurable as well.
\end{lemma}

\begin{theorem}
If $\mu$ and $\nu$ are two probability measures on $\Real^d$ and $\mu$ is atomless, then there exists at least a transport map $T$ such that $\Tmeasure{\mu}{T}=\nu$.
\end{theorem}